\section{UNIX system programming in ocaml}
\label{sec:unix-syst-progr}

\subsection{chap1}
\label{sec:chap1}

\begin{enumerate}
\item Modules Sys and Unix \\
  \textbf{Sys} containts those functions common to Unix and Windows.
  \textbf{Unix} contains everything specific to Unix.

  The \textit{Sys} and \textit{Unix} modules can override certain
  functions of the \textit{Pervasives} module
  \begin{alternate}
Unix.stdin;;
- : Batteries.Unix.file_descr = <abstr>
Pervasives.stdin;;
- : in_channel = <abstr>
\end{alternate}

\begin{ocamlcode}
  <prog.{native,byte}> : use_unix
  ocamlmktop -o ocamlunix unix.cma
\end{ocamlcode}

When running a program from a shell, the shell passes \textbf{arguments} and
\textbf{environment} to the program. When a program terminates
prematurely because \textit{an exception was raised but not caught}, it makes
an implicit call to \textit{exit 2}. For \textit{at\_exit}, the last
function to be registered is called first, and it can not be
unregistered. However, we can walk around it using global variables.

\begin{ocamlcode}
  Sys.argv, Sys.getenv , Unix.environment, 
  Pervasives.exit, Pervasives.at_exit, Unix.handle_unix_error
\end{ocamlcode}
\begin{alternate}
Sys.argv;;
\end{alternate}
\begin{ocamlcode}
- : string array =
[|"/Users/bob/SourceCode/ML/godi/bin/ocaml"; "dynlink.cma";
"camlp4of.cma"; "-warn-error"; "+a-4-6-27..29"|]
\end{ocamlcode}
\begin{alternate}
  Unix.environment ();;
\end{alternate}
\begin{ocamlcode}
- : string array =
[|"TERM=dumb"; "SHELL=/bin/bash";
  "TMPDIR=/var/folders/R4/R4awSXDIH6GpuuMmaVeCzU+++TI/-Tmp-/";
  "LIBRARY_PATH=/opt/local/lib/";
  "EMACSDATA=/Applications/Aquamacs.app/Contents/Resources/etc";
  "Apple_PubSub_Socket_Render=/tmp/launch-mcHkKo/Render";
  "EMACSPATH=/Applications/Aquamacs.app/Contents/MacOS/bin";
  "INCLUDE_PATH=/opt/local/include/"; "EMACS=t"; "USER=bob";
  "LD_LIBRARY_PATH=/opt/local/lib/"; "COMMAND_MODE=unix2003"; "TERMCAP=";
  "SSH_AUTH_SOCK=/tmp/launch-g9AcyQ/Listeners";
  "__CF_USER_TEXT_ENCODING=0x1F5:0:0"; "COLUMNS=68";
  "PATH=/opt/local/sbin:/usr/local/smlnj/bin:/usr/local/lib:/Applications/MATLAB_R2010b.app/bin:~/SourceCode/scala/scala-2.9.0.final/bin:/Users/bob/SourceCode/scripts:~/lib/emacs/customize:/usr/local/git/bin:/Users/bob/Racket/bin:/Users/bob/.cabal/bin:/Users/bob/SourceCode/ML/godi/bin:/Users/bob/SourceCode/ML/godi/sbin:/usr/texbin/:/bin:/usr/bin:/opt/local/bin/:/usr/local/lib/:/usr/local/bin/";
  "_=/usr/local/bin/ledit"; "C_INCLUDE_PATH=/opt/local/include/";
  "PWD=/Users/bob/SourceCode/Notes/ocaml-book";
  "TEXINPUTS=.:/Applications/Aquamacs.app/Contents/Resources/lisp/aquamacs/edit-modes/auctex/latex:";
  "EMACSLOADPATH=/Applications/Aquamacs.app/Contents/Resources/lisp:/Applications/Aquamacs.app/Contents/Resources/leim";
  "SHLVL=3"; "HOME=/Users/bob"; "LOGNAME=bob";
  "CAMLP4_EXAMPLE=/Users/bob/SourceCode/ML/godi/build/distfiles/ocaml-3.12.0/camlp4/examples/";
  "DISPLAY=/tmp/launch-sXEeNT/org.x:0"; "INSIDE_EMACS=23.3.50.1,comint";
  "EMACSDOC=/Applications/Aquamacs.app/Contents/Resources/etc";
  "SECURITYSESSIONID=616cd3"|]
\end{ocamlcode}

\item ERROR handling \\
  \begin{bluetext}
    exception Unix_error of error * string * string
    type error = E2BIG | ... |EUNKNOWERR of int 
  \end{bluetext}
  The second arg  of \textit{Unix\_error} is the name of the system
  call that raised the error, the third, if possible, identifies the
  object on which the error occured (i.e. file name).
  \textit{Unix.handle\_unix\_error}, if this raises the exception
  \textit{Unix\_error}, displays the message, and \textit{exit 2}


  \begin{ocamlcode}
let handle_unix_error2 f arg = let open Unix in 
  try
     f arg
  with Unix_error(err, fun_name, arg) ->
  prerr_string Sys.argv.(0);
  prerr_string ": \"";
  prerr_string fun_name;
  prerr_string "\" failed";
  if String.length arg > 0 then begin
     prerr_string " on \"";
     prerr_string arg;
     prerr_string "\"" end;
     prerr_string ": ";
     prerr_endline (error_message err);
     exit 2;;  
   \end{ocamlcode}
   
   \begin{bluetext}
val handle_unix_error2 : ('a -> 'b) -> 'a -> 'b = <fun>     
\end{bluetext}

\begin{bluetext}
  let rec restart_on_EINTR f x =
  try f x with Unix_error (EINTR, _, _) -> restart_on_EINTR f x  
\end{bluetext}

\begin{alternate}
finally;;
- : (unit -> unit) -> ('a -> 'b) -> 'a -> 'b = <fun>
finally (fun _ -> print_endline "finally") (fun _ -> failwith "haha") ();;
\end{alternate}
\begin{ocamlcode}
finally
Exception: Failure "haha".
\end{ocamlcode}

In case the program fails, i.e. raises an exception, \textit{the finalizer is
run and the exception  ex is raised again}. If \textbf{both} the main function
and the finalizer fail, the finalizer's exception is raised.
\end{enumerate}

\subsection{chap2}
\label{sec:chap2}

\begin{enumerate}
\item Files \\
  \textbf{File} covers \textit{standard files, directories, symbolic
    links, special files(devices), named pipes, sockets}
\item \textbf{Filename}  module \\
  makes filename cross platform
  \begin{bluetext}
    val current_dir_name : string
    val parent_dir_name : string
    val dir_sep : string
    val concat : string -> string -> string
    val is_relative : string -> bool
    val is_implicit : string -> bool
    val check_suffix : string -> string -> bool
    val chop_suffix : string -> string -> string
    val chop_extension : string -> string
    val basename : string -> string
    val dirname : string -> string
    val temp_file : ?temp_dir:string -> string -> string -> string
    val open_temp_file :
      ?mode:open_flag list ->
      ?temp_dir:string -> string -> string -> string * out_channel
    val temp_dir_name : string
    val quote : string -> string
  \end{bluetext}

  non-directory files can have \textbf{many parents}(we say that they have many
  \textbf{hard links}). There are also \textit{symbolic links} which
  can be seen as \textit{non-directory} files containing a path, conceptually,
  this path can be obtained by reading the contents of the symbolic
  link like an ordinary file. Whenever a symbolic link occurs in the
  \textbf{middle} of  a path, we have to follow its path
  transparently.
  \begin{bluetext}
    p/s/q -> l/q (l is absolute)
    p/s/q -> p/l/q (l is relative)
  \end{bluetext}
  \begin{bluetext}
    Sys.getcwd, Sys.chdir, Unix.chroot
  \end{bluetext}
  \textit{Unix.chroot p} makes the node p, which should be a
  directory, the root of the \textit{restricted} view of the
  hierarchy. Absolute paths are then interpreted according to this new
  root p (and .. at the new root is itself).
  Due to hard links, a file can have many different names.

\begin{ocamlcode}
Unix.(link, unlink,symlink,rename);;
\end{ocamlcode}
\begin{ocamlcode}
- : (string -> string -> unit) * (string -> unit) *
    (string -> string -> unit) * (string -> string -> unit)    
  \end{ocamlcode}
  
  \textit{unlink f} is like \textit{rm -f f}, \textit{link f1 f2} is
  like \textit{ln f1 f2}, \textit{symlink f1 f2} is like \textit{ln -s
  f1 f2}, rename f1 f2 is like \textit{mv f1 f2}

  A file descriptor represents a pointer to a file along with other
  information like the current read/write position in the file, the
  access rights, etc. \textbf{file\_descr}

  \begin{ocamlcode}
    Unix.(stdin,stdout,stderr);;
  \end{ocamlcode}
  
  \begin{ocamlcode}
  - : Batteries.Unix.file_descr * Batteries.Unix.file_descr *
    Batteries.Unix.file_descr    
  \end{ocamlcode}
  without redirections, the three descriptors refer to the terminal.
  \begin{bluetext}
    cmd > f ; cmd 2 > f
  \end{bluetext}
\item Meta attributes, types and permissions \\


  \begin{alternate}
Unix.(stat,lstat,fstat);;
  \end{alternate}
\begin{ocamlcode}  
  (string -> Batteries.Unix.stats) *
  (string -> Batteries.Unix.stats) *
  (Batteries.Unix.file_descr -> Batteries.Unix.stats)    
\end{ocamlcode}
  \textit{lstat} returns information about the symbolic link itself,
  while \textit{stat} returns information about the file that link
  points to.
  \begin{alternate}
Unix.(lstat &&& stat) "/usr/bin/al";;    
  \end{alternate}
  \begin{ocamlcode}
({Batteries.Unix.st_dev = 234881026; Batteries.Unix.st_ino = 843893;
  Batteries.Unix.st_kind = Batteries.Unix.S_LNK; (* link *)
  Batteries.Unix.st_perm = 493; Batteries.Unix.st_nlink = 1;
  Batteries.Unix.st_uid = 0; Batteries.Unix.st_gid = 0;
  Batteries.Unix.st_rdev = 0; Batteries.Unix.st_size = 46;
  (* pretty  small as a link *)
  Batteries.Unix.st_atime = 1273804908.;
  Batteries.Unix.st_mtime = 1273804908.;
  Batteries.Unix.st_ctime = 1273804908.},

 {Batteries.Unix.st_dev = 234881026; Batteries.Unix.st_ino = 840746;
  Batteries.Unix.st_kind = Batteries.Unix.S_REG; (*  regular file *)
  Batteries.Unix.st_perm = 493; Batteries.Unix.st_nlink = 1;
  Batteries.Unix.st_uid = 0; Batteries.Unix.st_gid = 80;
  Batteries.Unix.st_rdev = 0; Batteries.Unix.st_size = 163;
  (* maybe bigger *)
  Batteries.Unix.st_atime = 1323997427.;
  Batteries.Unix.st_mtime = 1271968805.;
  Batteries.Unix.st_ctime = 1273804911.})    
\end{ocamlcode}

  A file is uniquely identified by the pair made of its device
  number(typically the disk partition where it is located)
  \textit{st\_dev} and its inode number \textit{st\_ino}

  All the users and groups on the machine are usually described in the
  \textit{/etc/passwd, /etc/groups} files.
  \begin{bluetext}
    st_uid
    st_gid
    getpwnam, getgrnam, (by name, get passwd_entry, group_entry)
    getpwuid, getgrgid (by id)
    getlogin, getgroups
    chown, fchown
  \end{bluetext}

  \begin{ocamlcode}
Unix.getlogin () |> Unix.getpwnam;;    
\end{ocamlcode}

\begin{ocamlcode}
{Batteries.Unix.pw_name = "bob"; Batteries.Unix.pw_passwd = "********";
 Batteries.Unix.pw_uid = 501; Batteries.Unix.pw_gid = 20;
 Batteries.Unix.pw_gecos = "bobzhang"; Batteries.Unix.pw_dir = "/Users/bob";
 Batteries.Unix.pw_shell = "/bin/bash"}

\end{ocamlcode}

for access rights, executable, writable, readable by the user owner,
group owner, other users. For a directory, the executable permission
means the right to enter it, and read permission the right to list its
contents. The special bits do not have meaning unless the \textbf{x}
bit is set. The bit \textit{t} allows sub-directories to inherit the
permissions of the parent directory. On a directory, the bit
\textit{s} allows the use of the directory's \textit{uid} or
\textit{gid} rather than the user's to create directories. For an
executable file, the bit \textit{s} allows the chaning at executation
time of the user's effective identity or group with the system calls
\textit{setuid} and \textit{setgid}

\begin{alternate}
Unix.(setuid, getuid);;
- : (int -> unit) * (unit -> int) = (<fun>, <fun>)  
\end{alternate}

\item operations on directries \\
  only the kernel can write in directories(when files are
  created). Opening a directory in write mode is \textit{prohibited}.

  \begin{alternate}
Unix.(opendir,readdir,rewinddir,closedir);;    
\end{alternate}

\begin{ocamlcode}
- : (string -> Batteries.Unix.dir_handle) *
    (Batteries.Unix.dir_handle -> string) *
    (Batteries.Unix.dir_handle -> unit) * (Batteries.Unix.dir_handle -> unit)  
  \end{ocamlcode}

  \textit{rewinddir} repositions the descriptor at the \textbf{beginning} of
  the directory.

  \begin{ocamlcode}
    mkdir, rmdir
  \end{ocamlcode}
  We can only remove a directory that is \textbf{already empty}. It is
  thus necessary to first recursively empty the contents of the
  directory and then remove the directory.

  \begin{ocamlcode}
exception Hidden of exn 
(** add a tag to exn *)
let hide_exn f x = try f x with exn -> raise (Hidden exn)
(** strip the tag of exn *)
let reveal_exn f x = try f x with Hidden exn -> raise exn 
\end{ocamlcode}
\item  File manipulation \\

  \begin{alternate}
Unix.openfile;;    
\end{alternate}

\begin{ocamlcode}
- : string ->
    Batteries.Unix.open_flag list ->
    Batteries.Unix.file_perm -> Batteries.Unix.file_descr
  \end{ocamlcode}
  Most programs use \textit{0o666} means \textit{rw-rw-rw-}. with the default
  creation mask of \textit{0o022}, the file is thus created with the permission
  \textit{rw-r--r--}. With a more lenient mask of 0o002, the file is
  created with the permissions \textit{rw-rw-r--}. The third argument
  can be anything as \textit{O\_CREATE} is not specified.
  And to write to an empty file without caring any previous content,
  we use
  \begin{ocamlcode}
    Unix.openfile filename [O_WRONLY; O_TRUNC; O_CREAT] 0o666
  \end{ocamlcode}
  If the file is scripts, we create it with execution permission:
  \begin{ocamlcode}
    Unix.openfile filename [O_WRONLY; O_TRUNC; O_CREAT] 0o777
  \end{ocamlcode}
  If we want it to be confidential,
  \begin{ocamlcode}
    Unix.openfile filename [O_WRONLY; O_TRUNC; O_CREAT] 0o600
  \end{ocamlcode}
  The \textit{O\_NONBLOCK} flag guarantees that if the file is a named pipe or a
  special file then the file opening and subsequent reads and writes
  wil be non-blocking. The \textit{O\_NOCTYY} flag guarantees that if
  the file is a control terminal, it won't become the controlling
  terminal of the calling process.

  \begin{alternate}
    Unix.(read,single_write);;
  \end{alternate}
  \begin{ocamlcode}
  - : (Batteries.Unix.file_descr -> string -> int -> int -> int) *
    (Batteries.Unix.file_descr -> string -> int -> int -> int)
  \end{ocamlcode}    
  The \textit{string} hold the read bytes or the bytes to write. The 3rd
  argument is the start, the forth is the number.

  For writes, the number of bytes actually written is usually the
  number of bytes requested, with two exceptions
  (i) not possible to write (i.e. disk is full) (ii) the descript is a
  pipe or a socket open in non-blocking mode(async) (iii) due to
  OCaml, too large.

  The reason for (iii) is that internally OCaml uses auxiliary buffer
  whose size is bounded by a maximal value.

  OCaml also provides \textit{Unix.write} which iterates the writes
  until all the data is written or an error occurs. The problem is
  that in case of error there's no way to know the number of bytes
  that were \textit{actually written}. \textit{single\_write}
  preserves the atomicity of writes.

  For reads, when the current position is at the end of file, read
  returns zero. The convention \textit{zero equals end of file} also
  holds for special files, \textit{i.e. pipes and sockets}. For
  example, read on a terminal returns zero if we issue a
  \textit{Ctrl-D} on the input.

  But you may consider the blocking-mode in case.

  \begin{ocamlcode}
    Unix.close : file_descr -> unit 
  \end{ocamlcode}
  In contrast to Pervasives' channels, a file descriptor does not need
  to be closed to ensure that all pending writes have been performed
  as write requests are \textit{immediately} transmitted to the
  kernel. On the other hand, the number of descriptors allocated by a
  process is limited by the kernel(several hundreds to thousands).


  \begin{ocamlcode}
let buffer_size = 8192 
let buffer = String.create buffer_size 

(** this is unsatisfactory, if we copy an executable file, we would
like the copy to be also executable. *)
let file_copy input output = Unix.(
  let fd_in = openfile input [O_RDONLY] 0 in 
  let fd_out = openfile output [O_WRONLY; O_CREAT; O_TRUNC] 0o666 in 
  let rec copy_loop () = match read fd_in buffer 0 buffer_size with 
    |0 -> ()
    |r -> write fd_out buffer 0 r |> ignore; copy_loop () in 
  copy_loop ();
  close fd_in ; 
  close fd_out 
)


let copy () = 
  if Array.length Sys.argv = 3 then begin 
    file_copy Sys.argv.(1) Sys.argv.(2)
  end 
  else begin 
    prerr_endline 
      ("Usage: " ^ Sys.argv.(0) ^ "<input_file> <output_file>"); 
    exit 1 
  end 

let _  = Unix.handle_unix_error copy () 
\end{ocamlcode}

\begin{bluetext}
ocamlbuild find.byte -- find.ml find.xxxx    
\end{bluetext}

\begin{alternate}
ocamlbuild find.byte -- find.mlx find.xxxx
_build/find.byte: "open" failed on "find.mlx": No such file or directory
\end{alternate}
\item  system call \\
  For a system call, even if it does very little work, cost dearly --
  much more than a normal function call. So we need buffer to reduce
  the number of system call. For ocaml, the \textit{Pervasives} module
  adds another layer \textit{in\_channel, out\_channel}.

\item positioning and operations specific to certain file types

  \begin{alternate}
Unix.lseek;;
- : Batteries.Unix.file_descr -> int -> Batteries.Unix.seek_command -> int =
\end{alternate}

  File descriptors provide a uniform and media-independent interface
  for data communicatioin. However this uniformity breaks when we need
  to access all the features provided by a given media.

  For normal files, specific API
  \begin{ocamlcode}
Unix.(truncate,ftruncate);;
- : (string -> int -> unit) * (Batteries.Unix.file_descr -> int -> unit) =
\end{ocamlcode}
For symbolic links
\begin{ocamlcode}
Unix.(symlink, readlink);;
- : (string -> string -> unit) * (string -> string) = (<fun>, <fun>)  
\end{ocamlcode}

special files
\begin{enumerate}
\item /dev/null  black hole. (useful for ignoring the result)
\item /dev/tty* control terminals
\item /dev/pty* pseudo-terminals
\item /dev/hd* disks
\item /proc Under linux, system parameters organized as a file system.
\end{enumerate}

many special files ignore \textit{lseek}
\item terminals \\

  \begin{alternate}
Unix.(tcgetattr, tcsetattr);;
\end{alternate}
\begin{ocamlcode}
(Batteries.Unix.file_descr -> Batteries.Unix.terminal_io) *
(Batteries.Unix.file_descr ->
     Batteries.Unix.setattr_when -> Batteries.Unix.terminal_io -> unit)
\end{ocamlcode}
  
  \begin{alternate}
Unix.(tcgetattr stdout);;    
\end{alternate}

\begin{ocamlcode}
{Batteries.Unix.c_ignbrk = false; Batteries.Unix.c_brkint = true;
 Batteries.Unix.c_ignpar = false; Batteries.Unix.c_parmrk = false;
 Batteries.Unix.c_inpck = false; Batteries.Unix.c_istrip = false;
 Batteries.Unix.c_inlcr = false; Batteries.Unix.c_igncr = false;
 Batteries.Unix.c_icrnl = true; Batteries.Unix.c_ixon = false;
 Batteries.Unix.c_ixoff = false; Batteries.Unix.c_opost = true;
 Batteries.Unix.c_obaud = 9600; Batteries.Unix.c_ibaud = 9600;
 Batteries.Unix.c_csize = 8; Batteries.Unix.c_cstopb = 1;
 Batteries.Unix.c_cread = true; Batteries.Unix.c_parenb = false;
 Batteries.Unix.c_parodd = false; Batteries.Unix.c_hupcl = true;
 Batteries.Unix.c_clocal = false; Batteries.Unix.c_isig = false;
 Batteries.Unix.c_icanon = false; Batteries.Unix.c_noflsh = false;
 Batteries.Unix.c_echo = false; Batteries.Unix.c_echoe = true;
 Batteries.Unix.c_echok = false; Batteries.Unix.c_echonl = false;
 Batteries.Unix.c_vintr = '\003'; Batteries.Unix.c_vquit = '\028';
 Batteries.Unix.c_verase = '\255'; Batteries.Unix.c_vkill = '\255';
 Batteries.Unix.c_veof = '\004'; Batteries.Unix.c_veol = '\255';
 Batteries.Unix.c_vmin = 1; Batteries.Unix.c_vtime = 0;
 Batteries.Unix.c_vstart = '\017'; Batteries.Unix.c_vstop = '\019'}  
\end{ocamlcode}

it seems that ledit will change your input, and you can not get
\textit{Unix.(tcgetattr stdin)} work.

The code below works in real terminal, but does not work in
pseudo-terminals(like Emacs )

\begin{ocamlcode}
let read_passwd message = Unix.(
match 
   try 
    let default = tcgetattr stdin in 
    let silent = {default with c_echo = false; c_echoe = false ; 
                  c_echok = false; c_echonl = false ; } in 
     Some (default, silent)
   with _ -> None 
with
 |None -> Legacy.input_line Pervasives.stdin 
 |Some (default, silent) -> 
   print_string message ; 
   Legacy.flush Pervasives.stdout ; 
   tcsetattr stdin TCSANOW silent ; 
   try 
     let s = Legacy.input_line Pervasives.stdin in 
     tcsetattr stdin TCSANOW default; s 
   with x ->      tcsetattr stdin TCSANOW default; raise x 
    
);;
\end{ocamlcode}
 Sometimes a program needs to start another and connect its standard
 input to a terminal (or pseudo-terminal). To achieve that, we must
 manually look among the pseudo-terminals(/dev/tty[a-z][a-f0-9]) and
 find one that is not already open. We can open this file and start
 the program with this file on its standard input.

 The function \textit{tcsendbreak} sends an interrupt to the
 peripheral. The second argument is the duration of the interrupt.


 \begin{bluetext}
   tcdrain, tcflush, tcflow, setsid
 \end{bluetext}

\item locks on files
  \begin{bluetext}
Unix.lockf;;
- : Batteries.Unix.file_descr -> Batteries.Unix.lock_command -> int -> unit =    
\end{bluetext}

ocaml-expect
\begin{alternate}
let p = X.spawn "ocaml" [||];;
val p : X.t = <abstr>
X.expect p ~fmatches:[(fun s -> Some s)] [] "";;
- : string = "        Objective Caml version 3.12.1"
X.send p "3;;\n";;
- : unit = ()
X.expect p ~fmatches:[(fun s -> Some s)] [] "";;
- : string = "- : int = 3"  
\end{alternate}

not very powerful
\end{enumerate}

\subsection{chap3}
\label{sec:chap3}



%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../master"
%%% End: 
